---
title: 'Lab Report #2'
author: 'Lab Section C3: Yijia Liu, Stephanie Pan, Mohish Agrawal, Daniel Shen, Rashid Mammadov'
date: "10/2/2020"
output: word_document
---

```{r setup, message=FALSE, include=TRUE}
# Load packages
library(tidyverse) # includes dplyr and ggplot2
library(ggplot2)
library(hrbrthemes)
library(GGally)
library(weathermetrics)

# Load dataset
setwd("C:/Users/steph/Documents/Fall 2020 Semester/MA575 Linear Models/Lab/Group Project/Bike-Sharing-Dataset")
dailydat <- read.csv("day.csv",header=TRUE)

# Convert original temperature variable to Fahrenheit
dailydat$newtemp <- dailydat$temp*41
dailydat$tempfah <- celsius.to.fahrenheit(dailydat$newtemp, round = 2)
dailydat$yr <- as.factor(dailydat$yr)
levels(dailydat$yr) <- c("2011","2012")
attach(dailydat)
```

### Question 1
The relationship between weather patterns and the number of riders was assessed for this bike sharing project. In this report, the focus is on temperature, which is the covariate of interest (X) and the number of daily casual riders is our response (Y). The observations from both years, 2011 and 2012, are included in this analysis. 

### Question 2
The scatterplot of the data is provided below, with the daily casual riders on the Y-axis and temperature (Fahrenheit) on the X-axis. Visually, as temperature increases there is also a slightly higher number of daily casual bike riders for a given day. 
```{r , echo=TRUE}
ggplot(dailydat, aes(y=casual,x=tempfah)) + geom_point(size=1) + xlab("Temperature (Fahrenheit)") + ylab("Number of Daily Casual Riders")
```

### Question 3
The ordinary least squares (OLS) regression model was fitted using the daily casual number of riders as our response and temperature (in Fahrenheit) as our covariate.

```{r , echo=TRUE}
# Fit the OLS model for Y=casual riders, X=temperature (F) using all of the observations in our daily dataset
dailymodel <- lm(data=dailydat, casual~tempfah)
```

### Question 4
The summary output from the model is provided below. The coefficient estimates for the intercept is -1044.97 (95% CI -1261.77 to -828.16) and the estimate for the temperature is 27.61 (95%CI 24.51 to 30.72). In this instance if temperature (Fahrenheit) is 0, on average, the expected number of daily casual riders would be -1044.97; this would not be interpretable in the context of this setting since one cannot have a negative number of casual riders. For each degree increase, there would be on average an increase of 27.61 daily casual riders. Based on the critical t-value and corresponding p-value there is evidence at a significance level of 0.05 that temperature is positively associated with the number of daily casual riders (p<0.001).

```{r , echo=TRUE}
summary(dailymodel)
round(confint(dailymodel,level=0.95),6)
vcov(dailymodel)          #covariance matrix

```

### Question 5
The scatterplot of the response (Y) against the covariate (X) shows that there exists a positive relationship between temperature and daily casual rider count. The results of the OLS model provides evidence that the coefficient of the slope is not zero and suggests a positive linear relationship at a significance level of 0.05. A one-degree Fahrenheit increase in temperature is associated with increase of 27 daily casual riders, on average, at any given day. The scatterplot shows increasing variance as the temperature increase and this corresponds with the fact that the R-squared of our OLS model is 0.29. In other words, this means that only 29% of the variability of the given data is explained via our model that includes only temperature.

```{r , echo=TRUE}
dailydat$predlm <- predict(dailymodel,type="response")
ggplot(dailydat, aes(y=casual,x=tempfah)) + geom_point(size=1) +xlab("Temperature (Fahrenheit)") + 
  ylab("Number of Daily Casual Riders") + geom_line(aes(y=predlm), color="blue")
```

Even though this is a good initial model that aligns with our expectations – warmer temperature provides more incentive for casual riders to participate in the bike sharing program – this model can be improved with the inclusion of other factors that may explain more of the variability in the number of daily casual riders. In an exploratory analysis, stratifying by year (2011 and 2012) and applying the OLS models suggests other factors can strength the model and our findings.

```{r , echo=TRUE}
# Y vs X plot with linear regression fit including time as dummy variable
ggplot(dailydat, aes(x=tempfah, y=casual, color=yr)) + geom_point() + xlab("Temperature (F)") + ylab("Number of daily casual riders") + geom_smooth(method = "lm", se=FALSE) + facet_wrap(~ yr) + labs(title = "Scatterplot with OLS Stratified by Year")
```